---
output:
  github_document:
    html_preview: TRUE
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "## ",
  fig.path = "README-"
)
```

# caMST
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/caMST)](http://cran.r-project.org/package=caMST)
[![Travis-CI Build Status](http://travis-ci.org/AnthonyRaborn/caMST.svg?branch=master)](http://travis-ci.org/AnthonyRaborn/caMST)


## Installation
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("AnthonyRaborn/caMST") # the developmental version; no CRAN version yet
```

## Usage

Here are some examples demonstrating how to use this package for various adaptive test frameworks.

### Mixed Computerized Adaptive Multistage Testing (Mca-MST)

This is a method that follows the traditional multistage testing framework with one exception: the routing stage utilizes item-level adaptation to try to better estimate individual examinees' abilities before administering the additional stages. The current implementation allows for any three-stage design with a single routing module (e.g., 1-2-3, 1-5-3, etc.), though this will be extended for any arbitrary number of stages in future updates.

Here is an example with a 1-3-3 design that has 18 items, where each stage has the same number of items, maximum Fisher information is used for all item- and module-level adaptations, and the expected a posteriori estimator is used for the provisional and final ability estimates.

```{r loading functions, echo = FALSE}
source("R/mixed_adaptive_test_function.R")
source("R/module_selection_function.R")
source("R/routing_item_selection.R")
source("R/iterative_maximum_likelihood_theta_estimation.R")
```

```{r McaMST example}
# using simulated test data 
load("data/example_thetas.rda") # 5 simulated abilities
load("data/example_responses.rda") # 5 simulated responses
load("data/example_transition_matrix.rda") # the transition matrix for an 18 item 1-3-3 balanced design
load("data/cat_items.rda") # the items designated for use in the routing module with item-level adaptation
load("data/mst_items.rda") # the items designated for use in the second and third modules with module-level adaptation
load("data/example_module_items.rda") # the matrix specifying how the item data frame relates to the modules

# what does the data look like?
example_thetas
example_responses[,1:5] # first 5 items
example_transition_matrix
head(cat_items) # 564 items to choose from for the first module
head(mst_items) # 18 items in the three 2nd stage modules and another 18 in the three 3rd stage modules
head(example_module_items, 10) # notice that there are only 6 items in the first module!


# run the Mca-MST model
results <- mixed_adaptive_test(response_matrix = example_responses, 
                               cat_item_bank = cat_items, initial_theta = 0,
                               method = "EAP", item_method = "MFI", 
                               cat_length = 6, cbControl = NULL, cbGroup = NULL,
                               randomesque = 1, mst_item_bank = mst_items, 
                               modules = example_module_items, 
                               transition_matrix = example_transition_matrix)

# The function outputs a list with named elements; 
# each individual is his or her own element in the list.
# the first person's results:
results[[1]]

# How good was our estimate of the individual's abilities?
data.frame("True Theta" = example_thetas,
           "Estimated Theta" = unlist(lapply(results, '[[', 1)),
           "CI95 Lower Bound" = unlist(lapply(results, '[[', 1)) -
             1.96*unlist(lapply(results, '[[', 4)),
           "CI95 Upper Bound" = unlist(lapply(results, '[[', 1)) +
             1.96*unlist(lapply(results, '[[', 4)))
```


For these five individuals, the estimated ability level appears fairly close at worst and almost precisely correct at best. In all cases, the 95% confidence interval ( $\hat{\theta} \pm 1.96*SEM \end{equation}$ ) includes the simulated ability level, indicating that the Mca-MST method works well for these individuals.
