---
title: "Computerized Adaptive Testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computerized Adaptive Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A computerized adaptive test (CAT) selects items one at a time based on the
examinee's running ability estimate. This is the most flexible — and most
complex — testing framework, and it generally provides the most precise ability
estimates for a given test length.

## Running a CAT

The `computerized_adaptive_test()` function requires an item bank (in **catR**
format) and a response matrix.

```{r basic-cat}
library(caMST)

data(example_thetas)
data(example_responses)
data(cat_items)
```

Here we administer a short 3-item CAT using Maximum Fisher Information (MFI) for
item selection:

```{r short-cat}
catResults <- computerized_adaptive_test(
  cat_item_bank = cat_items,
  response_matrix = example_responses,
  randomesque = 1,
  maxItems = 3,
  nextItemControl = list(
    criterion = "MFI",
    priorDist = "norm",
    priorPar = c(0, 1),
    D = 1,
    range = c(-4, 4),
    parInt = c(-4, 4, 33),
    infoType = "Fisher",
    randomesque = 1,
    random.seed = NULL,
    rule = "precision",
    thr = .3,
    nAvailable = NULL,
    cbControl = NULL,
    cbGroup = NULL
  )
)
catResults
```

The result is an S4 object of class `CAT`. The `show` method prints a summary
including the average theta estimate, average SEM, and the average number of
items administered.

## Examining the results

The key slots of the `CAT` object are:

- `final.theta.estimate` — the final ability estimates for each examinee.
- `final.theta.SEM` — the corresponding standard errors of measurement.
- `final.items.seen` — which items each examinee was administered.
- `final.responses` — the responses to those items.

```{r cat-slots}
catResults@final.theta.estimate
catResults@final.theta.SEM
catResults@final.items.seen
```

## Comparing estimates to true abilities

```{r cat-comparison}
data.frame(
  True.Theta = example_thetas,
  Estimated.Theta = catResults@final.theta.estimate,
  SEM = catResults@final.theta.SEM
)
```

With only 3 items, the estimates are rough. A longer test generally improves
precision.

## A longer CAT with precision-based stopping

A more realistic CAT uses a longer maximum test length and may stop early if the
SEM falls below a threshold (here, 0.3). The `randomesque` parameter introduces
variability in item selection by choosing randomly from the top *k* most
informative items.

```{r longer-cat}
set.seed(12345)
catLong <- computerized_adaptive_test(
  cat_item_bank = cat_items,
  response_matrix = example_responses,
  randomesque = 5,
  maxItems = 18,
  nextItemControl = list(
    criterion = "MFI",
    priorDist = "norm",
    priorPar = c(0, 1),
    D = 1,
    range = c(-4, 4),
    parInt = c(-4, 4, 33),
    infoType = "Fisher",
    random.seed = NULL,
    rule = "precision",
    thr = .3,
    nAvailable = NULL,
    cbControl = NULL,
    cbGroup = NULL
  )
)
catLong
```

### Accuracy with a 95% confidence interval

```{r cat-ci}
data.frame(
  True.Theta = example_thetas,
  Estimated.Theta = catLong@final.theta.estimate,
  CI95.Lower = catLong@final.theta.estimate - 1.96 * catLong@final.theta.SEM,
  CI95.Upper = catLong@final.theta.estimate + 1.96 * catLong@final.theta.SEM
)
```

With up to 18 items and the precision stopping rule, each examinee's 95%
confidence interval contains the true theta.

## Key parameters

| Parameter | Purpose |
|---|---|
| `maxItems` | Maximum number of items to administer |
| `randomesque` | Size of the pool from which the next item is randomly drawn |
| `method` | Provisional theta estimation method (e.g., `"BM"`, `"EAP"`) |
| `final_theta_method` | Theta method for the final reported estimate (defaults to `method`) |
| `nextItemControl` | A list passed to `catR::nextItem` controlling item selection, stopping rules, and content balancing |

See `?computerized_adaptive_test` and the **catR** package documentation for
full details on each parameter.
