---
title: "Introduction to caMST"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to caMST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The **caMST** package provides functions for simulating and administering
computerized adaptive tests (CAT), multistage tests (MST), and mixed
computerized adaptive multistage tests (Mca-MST). This vignette introduces the
package's bundled data sets and the helper functions used to build and visualize
test designs.

## Included data

The package ships with several data sets used throughout the examples.

```{r data-overview}
library(caMST)

data(example_thetas)
example_thetas
```

`example_thetas` contains five simulated ability values on the theta scale. The
corresponding item responses are stored in `example_responses`, a data frame
with 5 rows (one per examinee) and 600 columns (one per item across all item
banks).

```{r responses}
data(example_responses)
dim(example_responses)
example_responses[, 1:6]
```

### Item banks

Three item banks are provided, each formatted for the **catR** package:

- `cat_items`: 564 items for the CAT routing stage (all assigned to module 1).
- `mst_items`: 36 items for the MST stages (modules 2--7), used alongside
  `cat_items` in `mixed_adaptive_test()`.
- `mst_only_items`: 42 items spanning modules 1--7, used for a standalone MST
  with `multistage_test()`.

```{r item-banks}
data(cat_items)
data(mst_items)
data(mst_only_items)

dim(cat_items)
head(cat_items)

dim(mst_items)
head(mst_items)

dim(mst_only_items)
head(mst_only_items)
```

### Transition and module matrices

The test design is described by two matrices:

- A **transition matrix** (`example_transition_matrix`) that specifies which
  modules can follow which. A `1` in row *i*, column *j* means an examinee who
  just completed module *i* may be routed to module *j*.
- A **module matrix** (`example_module_items`) that maps items to modules. A `1`
  in row *i*, column *j* means item *i* belongs to module *j*.

```{r design-matrices}
data(example_transition_matrix)
example_transition_matrix

data(example_module_items)
head(example_module_items, 10)
```

## Building design matrices with helper functions

Writing these matrices by hand is tedious and error-prone for larger designs.
The `generate_transition_matrix()` and `generate_modules()` functions build them
from simpler inputs.

### generate_transition_matrix()

Specify the total number of modules and a list describing, for each module,
which modules it can route to. Terminal modules (those with no further stage) use
`0`.

```{r generate-transition}
tm <- generate_transition_matrix(
  n_modules = 7,
  paths = list(
    "1" = c(2, 3, 4),
    "2" = c(5, 6),
    "3" = c(5, 6, 7),
    "4" = c(6, 7),
    "5" = 0,
    "6" = 0,
    "7" = 0
  )
)
tm
```

This reproduces `example_transition_matrix`:

```{r verify-transition}
identical(unname(tm), unname(example_transition_matrix))
```

### generate_modules()

When the item bank contains a `module` column, `generate_modules()` converts it
to the items-by-modules matrix that `multistage_test()` expects.

```{r generate-modules}
module_items <- generate_modules(mst_only_items, n_modules = 7)
head(module_items, 10)

identical(unname(module_items), unname(as.matrix(example_module_items)))
```

## Visualizing the test design

`transition_matrix_plot()` draws the routing structure so you can verify the
design at a glance.

```{r plot-transition, fig.width = 6, fig.height = 4}
transition_matrix_plot(example_transition_matrix, n_stages = 3)
```

The function also accepts the `MST` or `MAT` result objects directly, pulling
the transition matrix and stage count from the fitted object (see the MST and
Mca-MST vignettes for examples).

## Next steps

With the data and design matrices in hand, you can run any of the three test
types provided by **caMST**:

- `computerized_adaptive_test()` — pure item-level adaptation (see
  `vignette("computerized-adaptive-testing")`).
- `multistage_test()` — module-level adaptation, optionally with number-correct
  scoring (see `vignette("multistage-testing")`).
- `mixed_adaptive_test()` — a CAT routing stage followed by MST modules (see
  `vignette("mixed-adaptive-testing")`).
