---
title: "Mixed Adaptive Testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mixed Adaptive Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A mixed computerized adaptive multistage test (Mca-MST) combines the best of
both approaches: the first stage uses item-level adaptation (like a CAT) to
quickly estimate each examinee's ability, then subsequent stages use module-level
adaptation (like a traditional MST). This can produce more precise routing
decisions while retaining the content-control advantages of pre-assembled
modules.

## How it works

1. **CAT routing stage**: Items are selected one at a time from `cat_item_bank`
   using item-level adaptation (e.g., Maximum Fisher Information). The number of
   items is controlled by `cat_length`.
2. **MST stages**: The theta estimate from the routing stage seeds a standard
   multistage test. Modules from `mst_item_bank` are selected at the module
   level using `module_select` and the routing rules in `transition_matrix`.

The first of `n_stages` is the CAT routing stage, so a design with a CAT routing
stage followed by two MST stages uses `n_stages = 3`.

## Setting up the data

The Mca-MST needs two separate item banks — one for the CAT routing stage and
one for the MST stages — plus the standard design matrices.

```{r load-data}
library(caMST)

data(example_thetas)
data(example_responses)
data(example_transition_matrix)
data(cat_items)
data(mst_items)
data(example_module_items)
```

`cat_items` contains 564 items for the routing stage (all assigned to module 1),
while `mst_items` contains 36 items across modules 2--7:

```{r inspect-banks}
dim(cat_items)
head(cat_items)

dim(mst_items)
head(mst_items)
```

The module matrix and transition matrix describe only the MST portion of the
design:

```{r inspect-design}
example_transition_matrix
head(example_module_items, 10)
```

## Running a mixed adaptive test

```{r mat}
results <- mixed_adaptive_test(
  response_matrix = example_responses,
  cat_item_bank = cat_items,
  initial_theta = 0,
  method = "EAP",
  item_method = "MFI",
  cat_length = 6,
  cbControl = NULL,
  cbGroup = NULL,
  randomesque = 1,
  mst_item_bank = mst_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  n_stages = 3
)
results
```

## Examining the results

The `MAT` object contains the same slots as `MST`, plus the items selected
during the CAT routing stage:

```{r mat-theta}
results@final.theta.estimate
results@final.theta.SEM
```

### Accuracy with a 95% confidence interval

```{r mat-ci}
data.frame(
  True.Theta = example_thetas,
  Estimated.Theta = results@final.theta.estimate,
  CI95.Lower = results@final.theta.estimate - 1.96 * results@final.theta.SEM,
  CI95.Upper = results@final.theta.estimate + 1.96 * results@final.theta.SEM
)
```

For all five examinees, the 95% confidence interval contains the true ability
value.

### Items administered

The `final.items.seen` slot shows which items each examinee received. The first
`cat_length` columns come from the CAT routing stage; the remaining columns come
from the MST modules:

```{r items-seen}
results@final.items.seen
```

### Module routing paths

The `modules.seen` slot records which modules each examinee was routed through:

```{r modules-seen}
results@modules.seen
```

## Visualizing the design

`transition_matrix_plot()` works directly with `MAT` objects:

```{r mat-plot, fig.width = 6, fig.height = 4}
transition_matrix_plot(results)
```

## Key parameters

| Parameter | Purpose |
|---|---|
| `cat_length` | Number of items in the CAT routing stage |
| `item_method` | Item selection criterion for the routing stage (e.g., `"MFI"`) |
| `randomesque` | Pool size for randomized item selection in the routing stage |
| `module_select` | Module selection criterion for MST stages |
| `n_stages` | Total number of stages (CAT routing stage counts as stage 1) |
| `method` | Provisional theta estimation method for both stages |
| `final_theta_method` | Theta method for the final reported estimate |

See `?mixed_adaptive_test` for full documentation.
