---
title: "Multistage Testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multistage Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A multistage test (MST) administers pre-assembled modules of items, routing
examinees to the next module based on their performance. This is a practical
middle ground between a fixed-form test and a fully adaptive CAT: test
developers retain control over module content while still adapting to examinee
ability.

## Setting up the data

An MST requires four inputs:

1. An **item bank** with IRT parameters for every item.
2. A **module matrix** mapping items to modules.
3. A **transition matrix** describing the routing structure.
4. A **response matrix** with examinee responses.

```{r load-data}
library(caMST)

data(example_thetas)
data(example_responses)
data(example_transition_matrix)
data(mst_only_items)
data(example_module_items)
```

The example uses a 1-3-3 design: one routing module in the first stage, three
modules each in the second and third stages, for a total of 7 modules and 18
items per examinee.

```{r show-design}
example_transition_matrix
```

## Running a multistage test with MFI module selection

By default, `multistage_test()` uses Maximum Fisher Information (MFI) to choose
the most informative module at each transition.

```{r mst-mfi}
resultsMFI <- multistage_test(
  mst_item_bank = mst_only_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  method = "BM",
  response_matrix = example_responses,
  initial_theta = 0,
  model = NULL,
  n_stages = 3,
  test_length = 18
)
resultsMFI
```

### Examining the results

```{r mfi-results}
data.frame(
  True.Theta = example_thetas,
  Estimated.Theta = resultsMFI@final.theta.estimate,
  CI95.Lower = resultsMFI@final.theta.estimate -
    1.96 * resultsMFI@final.theta.SEM,
  CI95.Upper = resultsMFI@final.theta.estimate +
    1.96 * resultsMFI@final.theta.SEM
)
```

All five true ability values fall within their 95% confidence intervals.

## Alternative module selection: MLWMI

You can change the module selection criterion with the `module_select` argument.
Here we use the Maximum Likelihood Weighted Module Information (MLWMI) method:

```{r mst-mlwmi}
resultsMLWMI <- multistage_test(
  mst_item_bank = mst_only_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  method = "BM",
  response_matrix = example_responses,
  initial_theta = 0,
  model = NULL,
  n_stages = 3,
  module_select = "MLWMI",
  test_length = 18
)

data.frame(
  True.Theta = example_thetas,
  MFI.Theta = resultsMFI@final.theta.estimate,
  MLWMI.Theta = resultsMLWMI@final.theta.estimate
)
```

The two methods produce different routing paths and theta estimates, though both
recover the true abilities reasonably well with this small sample.

## Number-correct (NC) scoring

In applied testing settings, module routing sometimes uses raw number-correct
scores rather than IRT-based information. The `nc_list` argument supports this.

Each element of the list corresponds to a module that routes to other modules.
The values are the *maximum* cumulative number correct needed to be routed to
each downstream module. Terminal modules do not need entries.

For a 1-3-3 design where each stage has 6 items (18 total):

```{r nc-list}
nc_list <- list(
  module1 = c(4, 5, 7),
  module2 = c(8, 14, Inf),
  module3 = c(8, 14, 18),
  module4 = c(-Inf, 14, 18),
  method = "cumulative_sum"
)
```

`Inf` and `-Inf` prevent routing to unreachable modules in non-crossed designs
(e.g., an examinee routed to the easy second-stage module cannot reach the hard
third-stage module).

```{r mst-nc}
ncResults <- multistage_test(
  mst_item_bank = mst_only_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  method = "BM",
  response_matrix = example_responses,
  initial_theta = 0,
  model = NULL,
  n_stages = 3,
  test_length = 18,
  nc_list = nc_list
)
ncResults
```

Notice that the printed summary now identifies the NC scoring method used.

```{r nc-results}
data.frame(
  True.Theta = example_thetas,
  NC.Theta = ncResults@final.theta.estimate,
  MFI.Theta = resultsMFI@final.theta.estimate
)
```

With this example data, NC scoring produces comparable theta estimates.

## Visualizing the test design

`transition_matrix_plot()` accepts either a transition matrix or a fitted `MST`
object:

```{r mst-plot, fig.width = 6, fig.height = 4}
transition_matrix_plot(resultsMFI)
```

## Key parameters

| Parameter | Purpose |
|---|---|
| `n_stages` | Number of stages in the design |
| `test_length` | Total number of items each examinee answers |
| `module_select` | Module selection criterion: `"MFI"`, `"MLWMI"`, `"MPWMI"`, `"MKL"`, `"MKLP"`, or `"random"` |
| `nc_list` | Optional list for number-correct routing (see Details in `?multistage_test`) |
| `method` | Provisional theta estimation method |
| `final_theta_method` | Theta method for the final reported estimate |

See `?multistage_test` for full documentation.
