---
title: "Theta Trajectory Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Theta Trajectory Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

After administering a test, it is often useful to visualize how the provisional
ability estimate evolved over the course of the test. The
`theta_trajectory_plot()` function reconstructs the theta estimate at each
decision point and produces a **ggplot2** plot. An animated version is available
via `theta_trajectory_animate()`.

Both functions work with all four test types: `CAT`, `MST`, `MAT`, and `HAT`.

## Setup

```{r load}
library(caMST)
library(ggplot2)

data(example_thetas)
data(example_responses)
data(cat_items)
data(mst_only_items)
data(mst_items)
data(example_module_items)
data(example_transition_matrix)
```

## CAT trajectory

In a computerized adaptive test, the theta estimate updates after every item.
`theta_trajectory_plot()` connects these per-item estimates with a line.

```{r cat-run}
catResult <- computerized_adaptive_test(
  cat_item_bank = cat_items,
  response_matrix = example_responses,
  randomesque = 1,
  maxItems = 18,
  nextItemControl = list(
    criterion = "MFI",
    priorDist = "norm",
    priorPar = c(0, 1),
    D = 1,
    range = c(-4, 4),
    parInt = c(-4, 4, 33),
    infoType = "Fisher",
    random.seed = NULL,
    rule = "precision",
    thr = .3,
    nAvailable = NULL,
    cbControl = NULL,
    cbGroup = NULL
  )
)
```

```{r cat-plot}
theta_trajectory_plot(catResult, person = 1)
```

Each point shows the theta estimate and its standard error after incorporating
one more item. The error bars shrink as more items are administered and the
estimate becomes more precise.

### Wider confidence intervals

Use `ci_multiplier = 1.96` for approximate 95% intervals:

```{r cat-ci}
theta_trajectory_plot(catResult, person = 1, ci_multiplier = 1.96)
```

## MST trajectory

In a multistage test, the theta estimate updates once per module rather than
after every item. `theta_trajectory_plot()` uses `geom_step()` for MST segments
to reflect this: the estimate holds flat across items within a module and then
jumps when the next module is scored.

```{r mst-run}
mstResult <- multistage_test(
  mst_item_bank = mst_only_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  method = "BM",
  response_matrix = example_responses,
  initial_theta = 0,
  model = NULL,
  n_stages = 3,
  test_length = 18
)
```

```{r mst-plot}
theta_trajectory_plot(mstResult, person = 1)
```

The stage and module labels at the bottom of the plot show which module the
examinee was routed to at each stage.

## Mixed adaptive test (MAT) trajectory

A mixed adaptive test begins with a CAT routing stage (item-level adaptation),
then transitions to MST modules. The trajectory plot reflects both regimes:
per-item updates during the CAT stage, and per-module step functions during the
MST stages.

```{r mat-run}
matResult <- mixed_adaptive_test(
  response_matrix = example_responses[1:2, ],
  cat_item_bank = cat_items,
  initial_theta = 0,
  method = "EAP",
  item_method = "MFI",
  cat_length = 6,
  cbControl = NULL,
  cbGroup = NULL,
  randomesque = 1,
  mst_item_bank = mst_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  n_stages = 3
)
```

```{r mat-plot}
theta_trajectory_plot(matResult, person = 1)
```

The first portion of the trajectory (labeled "CAT") updates after every item.
The remaining stages switch to module-level adaptation with step functions.

## Hybrid adaptive test (HAT) trajectory

A hybrid adaptive test reverses the MAT order: MST modules come first, followed
by a CAT stage. The trajectory plot shows the MST step functions first, then
per-item CAT updates.

```{r hat-run}
cat_bank <- cat_items[!rownames(cat_items) %in% rownames(mst_only_items), ]

hatResult <- hybrid_adaptive_test(
  response_matrix = example_responses[1:2, ],
  cat_item_bank = cat_bank,
  mst_item_bank = mst_only_items,
  modules = example_module_items,
  transition_matrix = example_transition_matrix,
  n_stages = 4,
  cat_length = 6,
  initial_theta = 0,
  method = "EAP",
  item_method = "MFI",
  cbControl = NULL,
  cbGroup = NULL,
  randomesque = 1
)
```

```{r hat-plot}
theta_trajectory_plot(hatResult, person = 1)
```

## Comparing multiple respondents

### Overlaid on one panel

Pass a vector of person indices to overlay their trajectories. Each person is
distinguished by color.

```{r multi-overlay}
theta_trajectory_plot(catResult, person = c(1, 2, 3))
```

### Faceted into separate panels

Set `facet_person = TRUE` to give each person their own panel:

```{r multi-facet}
theta_trajectory_plot(catResult, person = c(1, 2, 3), facet_person = TRUE)
```

## Animated trajectories

`theta_trajectory_animate()` creates an animation that reveals the trajectory
one update at a time, using **gganimate**. This requires the **gganimate** and
**gifski** packages.

```{r animate, eval = FALSE}
# Preview in the RStudio viewer
theta_trajectory_animate(catResult, person = 1)

# Save to a GIF file
theta_trajectory_animate(catResult, person = 1, file = "cat_trajectory.gif")

# Adjust animation timing
theta_trajectory_animate(catResult, person = 1,
  transition_length = 1, state_length = 2, fps = 5)
```

The animation builds up the same plot shown by `theta_trajectory_plot()`,
adding one theta update per frame so the viewer can follow how the estimate
converges.

## Customizing the plot

Both functions return **ggplot2** objects, so you can add layers or modify the
theme as usual:

```{r customize}
theta_trajectory_plot(catResult, person = 1) +
  ggplot2::geom_hline(yintercept = example_thetas[1], linetype = "dashed",
                       color = "red") +
  ggplot2::labs(subtitle = "Dashed line = true theta") +
  ggplot2::theme_bw()
```
